\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2020-12-24 manuscript.Rnw
Updates to discussion and conclusion

Alternative approaches might be appropriate in other contexts. For example, if evenly spaced panel data is available, a multinomial logistic regressions can be used to predict transition probabilities. On the other hand, some users may prefer a functional approach that is arguably more human readable. To facilitate this type of approach, a \code{define_model()} block can be used to construct transition probability matrices using nonstandard evaluation. Each possible combination of the parameters and input data (treatment strategies and patients) is automatically constructed and the user can write custom expressions that create the transition probability matrices as a function of the parameters and input data. The same \code{define_model()} block may also be used to construct utility and cost models, if desired. Example that leverage multinomial logistic regression and \code{define_model()} are provided on the package website. The latter is used to reproduce two examples (a simple Markov model of an HIV treatment and a time-inhomogeneous Markov cohort model of total hip replacement) from the reference textbook by \citet{briggs2006decision}.

In Section~\ref{sec:ictstm-example}, we fit a clock-reset multi-state model and simulated outcomes with an individual-level simulation. An iCTSTM can, however, also be developed without fitting a statistical model using \proglang{R} by manually inputting the arguments (e.g., coefficients, probability distributions) of \code{params_surv} objects for each possible transition. If IPD is available for the reference treatment, \code{create_params()} can be use to obtain the baseline \code{params_surv} object from a fitted multi-state model, and relative treatment effects (e.g., hazard ratios) could be added as additional columns to the coefficients stored in the \code{params_surv} object. A similar approach was used in the partitioned survival analysis example above (Section~\ref{sec:psm-example}). A re-analysis of the total hip replacement model from the \citet{briggs2006decision} textbook using an individual-levels simulation is provided on the package website to illustrate.

@@

As mentioned in the introduction, \code{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling. It differs from \pkg{hesim} in that it focuses solely on cDTSTMs (i.e., Markov cohort models) and is not designed to facilitate integration of economic models with the statistical models used for parameter estimation. Another important difference is that \code{hesim} is considerably faster since all simulations are vectorized (i.e, performed at the \proglang{C++} level). Performance comparisons are provided on the package website. The time-inhomogeneous total hip replacement \citet{briggs2006decision} example was run with both packages using 1,000 PSA iterations. A cohort model in \pkg{heemod} completed in approximately 85 seconds while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \code{hesim} ran in 9 seconds.

@@

%% -- Conclusion ---------------------------------------------------------------
\section{Conclusion} \label{sec:conclusion}
\pkg{hesim} is a modular and computationally efficient \proglang{R} package for development and analysis of simulation models for economic evaluation. Discrete-time cohort and continuous-time individual-level models are supported, encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes as well as partitioned survival analysis. To facilitate integration of statistical methods and economic models and use of both patient- and aggregate-level data, parameters can be estimated by fitting a statistical model with \proglang{R}, inputting parameter estimates from an external source, or a combination of the two. Simulated costs and QALYs from a PSA can be used for for decision-analysis within a cost-effectiveness framework, allowing users to represent decision uncertainty (e.g., cost-effectiveness planes, cost-effectiveness acceptability curves, and cost-effectiveness acceptability frontiers), and conduct value of information analysis. Algorithms are efficiently coded and written in \proglang{C++} so that approaches that have been considered computationally intensive such as individual-level simulations and PSA can run very quickly. The modular design makes modeling more flexible and makes it straightforward to add new features.

>>>>

Alternative approaches might be appropriate in other contexts. For example, if evenly spaced panel data is available, a multinomial logistic regressions can be used to predict transition probabilities. On the other hand, some users may prefer a functional approach that is arguably more human readable. To facilitate this type of analysis, a \code{define_model()} block can be used to construct transition probability matrices using nonstandard evaluation. Each possible combination of the parameters and input data (treatment strategies and patients) is automatically constructed and the user can write custom expressions that create the transition probability matrices as a function of the parameters and input data. The same \code{define_model()} block may also be used to construct utility and cost models, if desired. Example that leverage multinomial logistic regression and \code{define_model()} are provided on the package website. The latter is used to reproduce two examples (a simple Markov cohort model of an HIV treatment and a time-inhomogeneous Markov cohort model of total hip replacement) from the reference textbook by \citet{briggs2006decision}.

In Section~\ref{sec:ictstm-example}, we fit a clock-reset multi-state model and simulated outcomes with an individual-level simulation. An iCTSTM can, however, also be developed without fitting a statistical model using \proglang{R} by manually inputting the arguments (e.g., coefficients, probability distributions) of \code{params_surv} objects for each possible transition. If IPD is available for the reference treatment, \code{create_params()} can be use to obtain the baseline \code{params_surv} object from a fitted multi-state model, and relative treatment effects (e.g., hazard ratios) could be added as additional columns to the coefficients stored in the \code{params_surv} object. A similar approach was used in the partitioned survival analysis example above (Section~\ref{sec:psm-example}). A re-analysis of the total hip replacement model from the \citet{briggs2006decision} textbook using an individual-level simulation is provided on the package website to illustrate.

@@

As mentioned in the introduction, \code{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling. It differs from \pkg{hesim} in that it focuses solely on cDTSTMs (i.e., Markov cohort models) and is not designed to facilitate integration of economic models with the statistical models used for parameter estimation. Another important difference is that \code{hesim} is considerably faster since all simulations are vectorized (i.e, performed at the \proglang{C++} level). Performance comparisons are provided on the package website. The time-inhomogeneous total hip replacement example from \citet{briggs2006decision} was run with both packages using 1,000 PSA iterations. A cohort model in \pkg{heemod} completed in approximately 85 seconds while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \code{hesim} ran in 9 seconds.

@@

%% -- Conclusion ---------------------------------------------------------------
\section{Conclusion} \label{sec:conclusion}
\pkg{hesim} is a modular and computationally efficient \proglang{R} package for development and analysis of simulation models for economic evaluation. Discrete time cohort and continuous time individual-level models are supported, encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes as well as partitioned survival analysis. To facilitate integration of statistical methods and economic models and use of both patient- and aggregate-level data, parameters can be estimated by fitting a statistical model with \proglang{R}, inputting parameter estimates from an external source, or a combination of the two. Simulated costs and QALYs from a PSA can be used for for decision-analysis within a cost-effectiveness framework, allowing users to represent decision uncertainty (e.g., cost-effectiveness planes, cost-effectiveness acceptability curves, and cost-effectiveness acceptability frontiers), and conduct value of information analysis. Algorithms are efficiently coded and written in \proglang{C++} so that approaches that have been considered computationally intensive such as individual-level simulations and PSA can run very quickly. The modular design makes modeling more flexible and it straightforward to add new features.
